---
title: "2022-11-05 Ukraine"
date-modified: "*Updated `r Sys.Date()`*"
format: 
  html:
    code-overflow: wrap
    theme: materia
editor_options: 
  chunk_output_type: console
---

30 Day Map Challenge for 2022. See [here](https://30daymapchallenge.com) for info.

## Ukraine (Day 5)

Using Ukraine data, we will explore the topographic roughness and river network.

### Libraries

```{r load-libraries}
suppressPackageStartupMessages({
  library(here);
  library(geodata);
  library(sf);
  library(rnaturalearth);
  library(tmap);
  library(dplyr);
  library(terra);
  library(fs);
  library(httr);
  library(viridis);
})
```

### Get Data

Grab elevation data for Ukraine and check coordinate ref system and resolution.

```{r get-dem-data}
#| echo: true
#| eval: false
ukr <- geodata::elevation_30s("UKR", path=here(tempdir()))
class(ukr) # check class
res(ukr)
cat(crs(ukr), "\n")
fs::dir_create("data_raw")
terra::writeRaster(ukr, file = "data_raw/ukr_elev_30s.tif")
```

```{r}
#| eval: true
#| echo: false
ukr <- rast(here("data_raw/ukr_elev_30s.tif"))

```

Make a quick plot:

```{r quick-plot}
#| error: false
#| message: false

plot(ukr, col=viridis::mako(96))
title("Ukraine DEM")

```

### Calculate Terrain Indices

First, let's calculate the terrain ruggedness index (*`TPI`*). A measure of the mean of the **absolute** differences between the values of a cell and the surrounding 8 cells. We also calculate aspect and flow direction.

```{r calc-asp}
# calc terrain ruggedness index: mean 
ukr_tri <- terrain(ukr, v = "TRI", neighbors=8)

# calc aspect 
ukr_asp <- terrain(ukr, v = "aspect", neighbors=8)

# calc flowdir
ukr_flowdir <- terrain(ukr, v = "flowdir", neighbors=8)
```

::: {layout-ncol=2}

```{r}
#| echo: false
#| fig-cap: "Topographic Roughness Index, palette reversed to show roughest areas in blue"
plot(ukr_tri, col=rev(viridis::cividis(300)))

```

```{r}
#| echo: false
#| fig-cap: "Flow Direction"
plot(ukr_flowdir, col=rev(viridis::mako(300)))
```

:::

### Get Rivers of Ukraine

Here we download from the `Hydro*` datasets found here: [https://www.hydrosheds.org/products/hydrorivers](https://www.hydrosheds.org/products/hydrorivers)

```{r get-rivers}
#| eval: false
# download rivers
library(httr)
res <- GET(url = "https://data.hydrosheds.org/file/HydroRIVERS/HydroRIVERS_v10_eu.gdb.zip",
           write_disk("data_raw/eu_rivers_v10.zip"),
           progress())
unzip(here("data_raw/eu_rivers_v10.zip"), exdir = "data")

```

```{r crop-rivers}
#| echo: true
#| eval: false

# check layers and crs
st_layers(here("data_raw/HydroRIVERS_v10_eu.gdb"))[1]
rivs <- st_read(here("data_raw/HydroRIVERS_v10_eu.gdb"), st_layers(here("data_raw/HydroRIVERS_v10_eu.gdb"))[1]$name) 
st_crs(rivs)

# crop to just ukr
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
ukr_sf <- world %>% filter(sov_a3=="UKR") %>% 
  st_transform(4326)
st_crs(ukr_sf)

# crop rivers by boundary
rivs_ukr <- rivs[ukr_sf,]

# fix geometry to s2
rivs_ukr$geometry <- rivs_ukr$Shape %>%
   s2::s2_rebuild() %>%
   sf::st_as_sfc()

# save out
fs::dir_create(here("data_clean"))
save(ukr_sf, rivs_ukr, file = here("data_clean/rivers_ukraine.rda"))
```

```{r load-dat}
#| eval: true
#| echo: false

load(here("data_clean/rivers_ukraine.rda"))
```


### Map It All

Now make a final map! Here we include the country border, rivers, and TRI.

::: {.column-page}

```{r tmap}
#| eval: true
#| echo: true
#| message: false
#| out-width: 110%

# make a col palette for rivers, darker for larger
blues <- rev(c('#08306b', '#08519c', '#2171b5', '#4292c6', '#6baed6', '#9ecae1', '#c6dbef', '#deebf7'))

library(tmap)

# map
tm1 <- tm_shape(ukr_tri) + 
  tm_raster(palette = "-cividis", alpha = 0.96, legend.show = TRUE) +
  tm_shape(ukr_sf) + 
  tm_sf(border.col = "gray",lwd = 10, border.lwd = 10, alpha=0)+  
  tm_shape(rivs_ukr) + tm_lines(col = "ORD_STRA", palette = blues, alpha=1, lwd = "ORD_STRA", legend.lwd.show = FALSE, legend.col.show=FALSE) +
  tm_credits(fontfamily = "Roboto Condensed", 
             text = "River data from HydroSHEDS database\n<http://www.hydrosheds.org> WWF (2006-2013)",position = "LEFT")+
  tm_layout(title.size = 0.5,
            main.title = "Rivers & Topographic Roughness in Ukraine",
            fontfamily = "Roboto Condensed")+
  tm_compass() + tm_scale_bar() +
  tm_layout(frame = FALSE, 
            legend.position = c("right","top"))
  
print(tm1)
fs::dir_create("figs")
tmap::tmap_save(tm1, filename = here("figs/20221105_ukraine_rivers.png"), width = 11, height = 8.5, dpi = 300)
```

:::
